---
title: "Homework 0 - Basic inference and GitHub submission"
author: "Chaewon Lee"
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: html_document
---

# Question 1

Given $X_i$ where $i = 1, ..., 10$, a random sample from a
distribution with cdf 

$$ F(x) = 1 - e^{-a x} $$

for $x \ge 0$ and zero otherwise, derive and calculate for the given
sample the maximum likelihood estimate (MLE) of $a$. 

```{r}
set.seed(1)
a <- 5
x <- rexp(10, rate=a)
```

$$ f(x) = ae^{-ax} $$

$$ L(a | X_1, ..., X_{10}) = f(X_1|a)*f(X_2|a)*...*f(X_{10}|a) \\ = ae^{-aX_1}*ae^{-aX_2}...*ae^{-aX_{10}} \\ = a^{10}e^{-a(X_1+X_2+...+X_10)}$$

We know that $\sum_{x = 1}^{10}x = 1.685244$
$$ L(a | X_1, ..., X_{10}) = a^{10}e^{-1.685244a} $$

Here, we take the natural logarithm.
$$ l(a | X_1, ..., X_{10}) = 10ln{(a)}-1.685244a $$
$$ \displaystyle \frac{\partial l(a | X_1, ..., X_{10})}{\partial a} = \frac{10}{a}-1.685244=0 $$
Therefore, MLE of $a = 5.933857\approx 5.9$.



# Question 2

Suppose we have a bivariate distribution with 

$$f(x, y) = x y^2 / 13 $$

with support $\{(1,1), (1,2), (2,2)\}$. Compute $E(X|Y=2)$

```{r echo=FALSE}
f <- function(x,y) x * y^2 / 13
feval <- c(f(1,1), f(1,2), f(2,2))
barplot(feval, names.arg=c("(1,1)","(1,2)","(2,2)"),
        ylim=c(0,1), xlab="(x,y)", ylab="density")
```

From the provided pdf of (x,y), we know that f(1,1) = $\frac{1}{13}$, f(1,2) = $\frac{4}{13}$, f(2,1) = 0, f(2,2) = $\frac{8}{13}$.
$$ E(X|Y=2) = \sum_{X=x} x f(x|Y=2) \\ = 1*\frac{f(x=1,y=2)}{f_{y}(y=2)} + 2*\frac{f(x=2,y=2)}{f_{y}(y=2)} \\ = 1*\frac{4/13}{12/13} + 2*\frac{8/13}{12/13} \\ = 1* \frac{1}{3} + 2*\frac{2}{3} \\ = \frac{5}{3} $$



# Question 3

There are three coins in a bag, two are fair with 
$P(\textrm{heads}) = 0.5$ and one is biased with 
$P(\textrm{heads}) = 0.75$. Suppose we pull a coin out and flip it 4
times, getting heads 3 times. What is the posterior probability the
coin was the biased one?

```{r}
coin <- c(1,1,0,1)
```

Let ð‘‹ be the random variable that indicates the number of flips that revealed head.

Here, $\textrm{P({H,H,T,H})} = P(X = 3)$.

Coin A and Coin B are fair coins with $\textrm{P(H|A) = P(H|B) = 0.5}$, and Coin C is a biased coin with P(H|C) = 0.75.

What we want to know is $\textrm{P(C|{H,H,T,H})=P(C|X=3})$.

And we know that $\textrm{P(A)=P(B)=P(C)}=\frac{1}{3}$.

Since each flip is independent, and following the Bayes' theorem,

$P(A,X=3)=P(X=3|A)*P(A)=P(H|A)*P(H|A)*P(T|A)*P(H|A)*P(A)=\frac{1}{3}*(0.5)^3*(0.5)$

As both coin A and B are fair coins, we can summarize as follows. 

$$P(A, X=3) = P(B, X=3) =  \frac{1}{3}*(0.5)^3*(0.5) = 0.02083333$$
$$P(C, X=3) = P(X=3|C)*P(C)= P(H|C)*P(H|C)*P(T|C)*P(H|C)*P(C)=\frac{1}{3}*(0.75)^3*(0.25)=0.1875$$
$$P(X=3) = P(A,X=3)+P(B,X=3)+P(C,X=3)$$

$$P(C|X=3) = \frac{P(C, X=3)}{P(A, X=3) + P(B, X=3) + P(C, X=3)}= \frac{0.1875}{0.02083333+0.02083333+0.1875}=0.818181 \approx 0.82 $$


